#Initialisation
---
title: "CAP_script"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GTJ)
library(sf)
library(logspline)
library(geosphere)
library(raster)
library(maxnet)
library(RStoolbox)
```

## Path setup
```{r}
lidarpath <- file.path("D:/Documents/gtj_local/data/lidar_data/full/DATA/")
speciespath <- file.path(lidarpath, "Species/CAP/")
rawdatapath <- file.path("D:/Documents/gtj_local/data/raw_data/")
scriptpath <- file.path("D:/Documents/gtj_local/R/irstea_git/vignettes/")
outputpath <- file.path(scriptpath, "output/")
```

##Variables setup
```{r}
study_period=c(2005:2018)
```

# Capercaillie data - observation distance
```{r}
#Distance data and logspline GDT
OBSGDTshp <- st_read(file.path(speciespath, "OBSGDTtot2018.shp"))
OBSGDTshp$distance=as.numeric(as.character(OBSGDTshp$distance))

Tracks_distance <- subset(OBSGDTshp, !OBSGDTshp$PROSPID=="NA" & OBSGDTshp$PROSPSA=="Hivernales" & OBSGDTshp$distance < 100)

fitlogspline1 <- logspline(Tracks_distance$distance, lbound = 0, ubound = 100)
distances <- seq(0, 100, len = 501)
hist(as.numeric(Tracks_distance$distance), freq = FALSE, 20)
lines(distances, dlogspline(distances, fitlogspline1), col = "red")
```

# Study area (Champfromier)
```{r}
limits_CAP <- st_read(file.path(lidarpath, "Env_var/sites/Champfromier.geojson"))
```



#JURA tracks
```{r}
TracksJURA <- st_read(file.path(speciespath, 'IndividualLines2018_JURA.shp'))
TracksJURA <- subset(TracksJURA, SAISON=="Hivernales" &  Year %in% study_period)
TracksID_JURA <- TracksJURA$PROSPID
```

#JURA obs
```{r}
OBSGDTshp_JURA <- subset(OBSGDTshp, PROSPID %in% TracksID_JURA & Saison=="hiver" & PROSPSA=="Hivernales" & distance < 100 & Year %in% study_period)
plot(st_geometry(OBSGDTshp_JURA))
plot(st_geometry(OBSGDTshp), col = "red", add = TRUE)
plot(st_geometry(limits_CAP), add = TRUE, border = "black")
```


#AIN tracks
```{r}
TracksAIN <- st_read(file.path(speciespath, 'IndividualLines2018_AIN_GESTION.shp'))
TracksAIN <- subset(TracksAIN, SAISON=="Hivernales" &  Year %in% study_period)
TracksID_AIN <- TracksAIN$PROSPID
```

#AIN obs
```{r}
OBSGDTshp_AIN <- subset(OBSGDTshp, PROSPID %in% TracksID_AIN & Saison=="hiver" & PROSPSA=="Hivernales" & distance < 100 & Year %in% study_period)
plot(st_geometry(OBSGDTshp_AIN))
plot(st_geometry(OBSGDTshp), col = "red", add = TRUE)
plot(st_geometry(limits_CAP), add = TRUE, border = "black")
```

#Importing LiDAR outlines
```{r}
Lidar_AIN_contour <- st_read(file.path(speciespath, "Aire_gestion_CAP_AIN_buffer.shp"))
Lidar_JURA_contour <- st_read(file.path(speciespath, "Aire_gestion_CAP_JURA_buffer.shp"))
```

#Plot differences between JURA and AIN regions
```{r}
# plot(st_geometry(OBSGDTshp))
par(mfrow = c(1, 2))
plot(st_geometry(OBSGDTshp_JURA), col = "blue", axes = TRUE, main = "JURA")
plot(st_geometry(Lidar_JURA_contour), add = TRUE, border = "green")
plot(st_geometry(limits_CAP), add = TRUE, border = "red")

plot(st_geometry(OBSGDTshp_AIN), col = "green", axes = TRUE, main = "AIN")
plot(st_geometry(Lidar_AIN_contour), add = TRUE, border = "blue")
plot(st_geometry(limits_CAP), add = TRUE, border = "red")

```



#Setting up the predictions
```{r}
target_path_15_8ha_Mean <- file.path(lidarpath, "Env_var/MW1_8ha/Mean", "/")
target_path_15_8ha_sd <- file.path(lidarpath, "Env_var/MW1_8ha/sd/", "/")
new_env_raster <- raster::stack(
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_H.nb2_5relative_density.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_H.nb10_20relative_density.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_H.nb20_30relative_density.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_H.nb0.5_1ratio.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_H.nb2_5ratio.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_H.nb5_10ratio.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_H.simpson.tif"),
  paste0(target_path_15_8ha_Mean, "Distance_MV_1_8ha_MeanskiAIN.tif"),
  paste0(target_path_15_8ha_Mean, "Distance_MV_1_8ha_MeanRoadAIN_IGN.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_H.nb0.5_1relative_density.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_H.nb2_5relative_density.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_H.nb10_20relative_density.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_H.nb20_30relative_density.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_H.nb0.5_1ratio.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_H.nb2_5ratio.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_H.nb5_10ratio.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_H.simpson.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_Treeinf10.density.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_Tree10.density.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_Tree20.density.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_Tree.giniH.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_G.s1000_Inf.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_G.s200_1000.tif"),
  paste0(target_path_15_8ha_Mean,"AINraster_MV_1_8ha_Meancor_G.s20_200.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_Treeinf10.density.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_Tree10.density.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_Tree20.density.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_Tree.giniH.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_G.s1000_Inf.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_G.s200_1000.tif"),
  paste0(target_path_15_8ha_sd,"AINraster_MV_1_8ha_sdcor_G.s20_200.tif"),
  paste0(target_path_15_8ha_Mean, "AINraster_MV_1_8ha_Meancor_Tree.sumS.tif"),
  paste0(target_path_15_8ha_sd, "AINraster_MV_1_8ha_sdcor_Tree.sumS.tif"),
  paste0(target_path_15_8ha_Mean, "raster_MV_1_8ha_Mean_propConifG2.tif"),
  paste0(target_path_15_8ha_sd, "raster_MV_1_8ha_sd_propConifG2.tif"),
  paste0(target_path_15_8ha_Mean, "bati_surfaceratio_1_8ha_tot2.tif"), quick=TRUE
)

names(new_env_raster) <- c( "Canopy0205",  "Canopy1020", "Canopy2030","penetrationratio0501", "penetrationratio0205","penetrationratio0510","Simpson", "Ski", "Road",  "Canopy0501sd", "Canopy0205sd", "Canopy1020sd", "Canopy2030sd", "penetrationratio0501sd", "penetrationratio0205sd","penetrationratio0510sd","Simpsonsd", "Tree_densityinf10", "Tree_density10", "Tree_density20", "Tree_gini", "grassland", "middle_gap", "small_gap",  "Tree_densityinf10sd", "Tree_density10sd", "Tree_density20sd", "Tree_ginisd", "grasslandsd", "middle_gapsd", "small_gapsd", "Tree_sumS", "Tree_sumSsd", "Prop_resG", "Prop_resGsd", "Bat")

predictors <- c("Canopy2030sd",  "penetrationratio0205",  "penetrationratio0501sd", "penetrationratio0205sd",  "Tree_densityinf10", "Tree_density10",  "Tree_gini",  "middle_gap",   "Prop_resG", "Prop_resGsd", "Bat" , "Road" ,"grassland" )
new_env_raster <- subset(new_env_raster, predictors)
```

#Clipping the rasters with the limits of the study area
```{r}
new_env_raster_champfromier <- crop(new_env_raster, extent(limits_CAP))
new_env_raster_champfromier <- mask(new_env_raster_champfromier, limits_CAP)
```

#Plotting all the predictors
```{r, fig.width=20, fig.height=20}
plot(new_env_raster_champfromier)
```

#Plotting one predictor, combined with tracks
```{r, fig.width=10, fig.height=10}
predictor_to_test = "Tree_density10"
plot(raster(new_env_raster_champfromier, predictor_to_test), main = predictor_to_test)
plot(st_geometry(limits_CAP), add = TRUE, border = "black", lwd=2)
plot(st_geometry(TracksAIN), add = TRUE)
#NE FONCTIONNE PAS :
plot(st_geometry(OBSGDTshp_AIN), add = TRUE, col = "red", lwd=20) 
plot(st_geometry(OBSGDTshp_JURA), add = TRUE, col = "red", lwd=20)
```


 
#Model data 
```{r}
#Presence dataframe creation
pres_pointsGDT <- data.frame(st_coordinates(OBSGDTshp), Pres = 1)

#Absence dataframe creation
TracksSP <- as(TracksAIN, Class = "Spatial")
backgroundGDT <- RANDOM_POINTS(TracksSP, n=10000, logplineobj = fitlogspline1, type="random")
backgroundGDT <- data.frame(X=backgroundGDT$x0, Y=backgroundGDT$y0, Pres=0)

#Merge dataframes
pres_bgGDT <- rbind(pres_pointsGDT, backgroundGDT) 

#Spatial point dataframe
coordinates(pres_bgGDT) <- ~X + Y
proj4string(pres_bgGDT) <- st_crs(OBSGDTshp)$proj4string

#Extraction of raster values corresponding to points in the dataframe
pres_bgGDT@data <- cbind(pres_bgGDT@data, extract(new_env_raster_champfromier, pres_bgGDT))

#Dataframe setup
pres_bgGDT <- as(pres_bgGDT, "data.frame")
pres_bgGDT <- na.omit(pres_bgGDT)

Pres_data <- as.vector(pres_bgGDT$Pres)
```

#Model fitting
```{r}
predictor_maxnet <- pres_bgGDT[, c("Canopy2030sd",  "penetrationratio0205",  "penetrationratio0501sd", "penetrationratio0205sd",  "Tree_densityinf10", "Tree_density10",  "Tree_gini",  "middle_gap",   "Prop_resG", "Prop_resGsd", "Bat" , "Road", "grassland" )]
#Add a product term between Canopy2030 and grassland

FORMULA <- maxnet.formula(Pres_data, predictor_maxnet, classes = "lq")
FORMULA <- paste0(FORMULA, "+ Canopy2030sd:grassland")
FORMULA <- as.formula(FORMULA)

model_maxnet <- maxnet(Pres_data, predictor_maxnet, f = FORMULA, regmult = 1)
```

#Map predictions
```{r}
#use it to predict results
predictor <- rasterToPoints(new_env_raster_champfromier)
predictor <- na.omit(predictor)
predictor <- as.data.frame(predictor)

predictor$predicted <- predict(model_maxnet, predictor,type=c("exponential"), clamp=FALSE)[, 1]
# summary(predictor[, c("x","y","predicted")])

maps_predicted <- raster::rasterFromXYZ(predictor[, c("x","y","predicted")], res=c(25,25))
crs(maps_predicted) <- "+proj=lcc +lat_1=49 +lat_2=44 +lat_0=46.5 +lon_0=3 +x_0=700000 +y_0=6600000 +ellps=GRS80 +units=m +no_defs"
plot(maps_predicted)

writeRaster(maps_predicted, paste0(speciespath, "/AIN_model_14_exponential.tif"), overwrite = TRUE, NAflag =  -9999, options=c("COMPRESS=LZW"))

#cloglog

predictor$predicted<-predict(model_maxnet, predictor,type=c("cloglog"), clamp=FALSE)[,1]

maps_predicted<-rasterFromXYZ(predictor[, c("x","y","predicted")], res=c(25,25))

#Save maps
crs(maps_predicted) <- "+proj=lcc +lat_1=49 +lat_2=44 +lat_0=46.5 +lon_0=3 +x_0=700000 +y_0=6600000 +ellps=GRS80 +units=m +no_defs"
writeRaster(maps_predicted, paste0(output_path_15_8ha, "AIN_model_14_cloglog.tif"), overwrite = TRUE, NAflag =  -9999, options=c("COMPRESS=LZW"))
```
